Data Management for LSST Special Programs

### 2023-05-20 sec_proc.tex
Update sec_proc.tex

It depends on the science goals related to the Special Program, but some 
Special Programs' data can and should be processed and served along with 
(or as a part of) the WFD data products (Section~\ref{ssec:proc_wfd}).
Others will require unique and separate data products 
(Section~\ref{ssec:proc_rdm}).

Cases where the Special Program obtains non-standard visits that cannot 
be processed by the DMS (Section~\ref{ssec:proc_bounds}), and/or where 
development of new algorithms or the allocation of significant additional 
computational resources are needed, will require user processing 
(Section~\ref{ssec:proc_user}).

@@

However some Special Programs are likely to use non-standard visits with 
shorter exposures (5 sec), and some are likely to acquire images with a 
significantly brighter sky background (e.g., twilight images) than 
nighttime survey images.

@@

None of the currently proposed Special Programs (\citedsp{PSTN-055}) 
seem likely to approach (or violate) any harware boundaries 
(Section~\ref{ssec:proc_bounds_hardware}).

@@

\subsection{RDM Including Special Programs in the WFD Data Products}\label{ssec:proc_wfd}

@@

This would include all standard and alternative visits in sky regions for which a template image exists.

This might also include some non-standard visits (shorter or longer exposures) as long as they can be processed by the Prompt 
pipeline\footnote{If a Special Program's science goals do require 
specialized templates and 
Prompt processing, the DMS will have the capability to load and use an 
alternative template for some regions, based on the image metadata. 
However, there would not be enough memory to hold alternative templates 
for the whole sky region.}.

@@

\textbf{Two potential issues with Prompt processing for DDFs}
(or any mini-survey that obtains a sequence of images without slewing)
are:

(1) As alert packets contain the full records of all associated 
{\tt DIASources} from the past 12 months, alerts for {\tt DIAObjects} in 
the DDFs might become very large.

(2) The association of {\it new} {\tt DIASources} into {\tt DIAObjects} 
will be somewhat compromised for a DDF sequence.
The processing for the second image begins when the processing for the 
first image is only halfway done; when the {\tt DIAObject} catalog has 
not yet updated with the new {\tt DIASources} detected in the first image.
Thus, the {\tt DIASource} from images one and two for a new transient 
would not be associated with a single {\tt DIAObject}, but instead would 
each instantiate a new {\tt DIAObject}.

@@

Since Solar System Processing takes \texttt{DIASources} as input, so any 
Special Programs images that can be run through the Alert Pipeline can 
also be incorporated into Solar System Processing.

@@

\textbf{Deep coadds and the Object catalog -- }
Some of the core LSST science goals require a WFD main survey data 
products of nearly {\it uniform depth}.
Whether and how to include any Special Programs data in the deep image 
coadds and the corresponding {\tt Object} catalog is left entirely to the 
discretion of the appropriate Rubin teams in Operations.

@@

For stacked images and their associated catalogs, the region labels should be
included but the program labels would not be needed (or possible).

@@

\textbf{Custom coadds -- }
As with the annual WFD deeply coadded images and associated catalogs, RMD would 
only generate one coadd (including all images to date) for a given special region.
Custom coadds (e.g., weekly, monthly) for specific science goals are considered 
user-generated data products, for both WFD and Special Programs.

Decisions about when to produce separate Special Programs data products are 
left to the discretion of RDM.

@@

For example, RDM-generated ``unique and separate" DDF data products might 
include:
\begin{itemize}
\item nightly-coadded images (24 h)
\item nightly-coadded difference images (24 h)
\item {\tt DIASource}- and {\tt DIAObject}-like catalogs for the nightly-coadds (24 h)
\item deeply-coadded images (all images to date; yearly)
\item templates for the nightly-coadded difference images (yearly)
\item {\tt Source}- and {\tt Object}-like catalogs for the nightly-coadded and deeply-coadded images (yearly)
\end{itemize}

@@

If either of the static short-exposure mini-surveys is executed, and if 
RDM finds that the LSST Science Pipelines can be reconfigured and used 
for short exposures, then the RDM-generated data products would likely be 
of a similar format to the Data Release data products 
(Section 4, \citeds{lse-163}).
For example, a separate repository of processed visit images and coadded 
images, and separate (but joinable) {\tt Source}, {\tt Forced Source}, and 
{\tt Object} catalogs.

@@

In cases where the observations are all obtained with a short season 
(e.g., a week spent observing a specific region), such ``annual" 
data products might be generated on an intermediate time scale, 
if resources allow.

@@

\textbf{Prompt Processing / Alert Production -- }
As the latency on processed visit image availability has an 80-hour
embargo, no user-generated pipeline will be able to process Special Programs 
data on a timescale similar to prompt processing and alert production 
(60 seconds to 24 hours).
No user-generated pipeline may contribute alerts to the LSST alert stream on 
any timescale.

>>>>

It depends on the science goals related to the Special Program, but some 
Special Programs' data can and should be processed and served along with 
(or as a part of) the WFD data products (Section~\ref{ssec:proc_wfd}).
Others will require RDM to generate unique and separate data products 
(Section~\ref{ssec:proc_rdm}).

Cases where the Special Program obtains non-standard visits that cannot 
be processed by the DMS (Section~\ref{ssec:proc_bounds}), and/or where 
development of new algorithms or the allocation of significant additional 
computational resources are needed, will require users tp generate unique
and separate data products (Section~\ref{ssec:proc_user}).

@@

However some Special Programs are likely to use non-standard visits with 
shorter exposures (5 sec), and some are likely to acquire images with a 
significantly brighter sky background (e.g., twilight images) than 
standard (or altnerative) WFD main survey visits.

@@

None of the currently proposed Special Programs listed in the Phase 2 
SCOC recommendations \citedsp{PSTN-055} 
seem likely to approach (or violate) any harware boundaries 
(Section~\ref{ssec:proc_bounds_hardware}).

@@

\subsection{Inclusion in the WFD Data Products}\label{ssec:proc_wfd}

@@

This would include all standard and alternative visits in sky regions for which 
a template image exists.
This might also include some non-standard visits (shorter or longer exposures) 
as long as they can be processed by the Prompt 
pipeline\footnote{If a Special Program's science goals do require 
specialized templates and 
Prompt processing, the DMS will have the capability to load and use an 
alternative template for some regions, based on the image metadata. 
However, there would not be enough memory to hold alternative templates 
for the whole sky region.}.

@@

\textbf{Two potential issues with Prompt processing for DDFs}, 
or any mini-survey that obtains a sequence of images without slewing, 
include:

(1) As alert packets contain the full records of all associated 
{\tt DiaSources} from the past 12 months, alerts for {\tt DiaObjects} in 
the DDFs might become very large.

(2) The association of {\it new} {\tt DiaSources} into {\tt DiaObjects} 
will be somewhat compromised for a DDF sequence.
The processing for the second image begins when the processing for the 
first image is only halfway done; when the {\tt DiaObject} catalog has 
not yet updated with the new {\tt DiaSources} detected in the first image.
Thus, the {\tt DiaSource} from images one and two for a new transient 
would not be associated with a single {\tt DiaObject}, but instead would 
each instantiate a new {\tt DiaObject}.

@@

Since Solar System Processing takes {\tt DiaSource}s as input, so any 
Special Programs images that can be run through the Alert Pipeline can 
also be incorporated into Solar System Processing.

@@

\textbf{Deep coadds and the {\tt Object} catalog -- }
Some of the core LSST science goals require a WFD main survey data 
products of nearly uniform depth.
Whether and how to include any Special Programs data in the deep image 
coadds and the corresponding {\tt Object} catalog is left entirely to the 
discretion of the RDM team in Operations.

@@

For stacked images and their associated catalogs, the region labels should be
included (but the program labels would not be needed or possible).

@@

\textbf{A note about intermediate-depth ``custom" coadds -- }
As with the annual WFD deeply coadded images and associated catalogs, RDM would 
only generate one coadd (including all images to date) for a given special region.
Custom coadds (e.g., weekly, monthly) for very specific science goals are considered 
user-generated data products, for both WFD and Special Programs.

Decisions about when to produce separate Special Programs data products are 
left to the discretion of RDM.
Appendix~\ref{sec:spcs} provides more detailed, step-by-step data processing 
examples for some potential Special Programs as further illustration.

@@

For example, RDM-generated ``unique and separate" DDF data products might 
include:
\begin{itemize}
\item nightly-coadded images (24 h)
\item nightly-coadded difference images (24 h)
\item {\tt DiaSource}- and {\tt DiaObject}-like catalogs for the nightly-coadds (24 h)
\item deeply-coadded images (all images to date; yearly)
\item templates for the nightly-coadded difference images (yearly)
\item {\tt Source}- and {\tt Object}-like catalogs for the nightly-coadded and deeply-coadded images (yearly)
\end{itemize}

@@

If either of the static short-exposure mini-surveys is executed, and if 
RDM finds that the LSST Science Pipelines can be reconfigured and used 
for short exposures, then the RDM-generated data products would likely be 
of a similar format to the Data Release data products 
(Section 4, \citeds{lse-163}).
For example, a separate repository of processed visit images and coadded 
images, and separate (but joinable) {\tt Source}, {\tt ForcedSource}, and 
{\tt Object} catalogs.

@@

In cases where the observations are all obtained with a short season 
(e.g., a week spent observing a specific region), such ``annual" 
data products might be generated and released on an intermediate time scale, 
if resources allow.

@@

\textbf{A note prompt processing and alert production -- }
As the latency on processed visit image availability has an 80-hour
embargo, no user-generated pipeline will be able to process Special Programs 
data on a timescale similar to prompt processing and alert production 
(60 seconds to 24 hours).
No user-generated pipeline may contribute alerts to the LSST alert stream on 
any timescale.
